SULI Report!

### 2015-08-20 doc/suli-report-swe.tex
Revised the abstract for the SULI report. #81

\begin{abstract}
As dark matter does not absorb or emit light, its distribution in the universe must be inferred through indirect effects such as the gravitational lensing of distant galaxies. While most sources are only weakly lensed, the systematic alignment of background galaxies around a foreground lens can constrain the mass of the lens which is largely in the form of dark matter. In this paper, a simple halo model is used to approximate the dark matter mass distribution of the Millennium Simulation on the galaxy and cluster scales. Background sources are generated at a redshift of ${z = 1.3857}$ and then weakly lensed by the foreground halos contained in a light cone centered along each line of sight, which is carried out by the publicaly available \texttt{Pangloss} framework. The modelpredicted ellipticities of the background galaxies after lensing are compared to the ``true'' weakly-lensed ellipticities determined by ray-traced convergence and shear from the simulation. Using the ellipticity-ellipticity and galaxy-mass correlation functions, I find that the \texttt{Pangloss} framework systematically underpredicts the shear power in both statistics and does not accurately capture the effect of dark matter structure at scales larger than 1 arcminute. Physical and computational shortcomings of the \texttt{Pangloss} framework are discussed, as well as potential improvements for upcoming work.
\end{abstract}

@@

%% Ellipticity-Ellipticity Correlation
\begin{figure}
    \centering
    \begin{subfigure}[h]{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/gg_corr.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The ellipticity-ellipticity correlation function for 3,240 sources at separation distances from 0.1 arcminutes to 2 arcminutes and a lightcone radius of $R=2$ arcminutes. While the correlation of the intrinsic ellitpicities before lensing (blue) shows no correlation, the correlation of the ray-traced ellipticities (green) and halo model predicted ellipticities (purple) both show significant positive correlation on scales up to 1 arcminute. The halo model predicts no correlation above this scale, while the ray-tracing shows positive correlation on all scales.}
        %The correlation of the intrinsic ellitpicities before lensing is plotted in blue, and shows no correlation. The correlation of the true ray-traced ellipticities is plotted in green and shows significant positive correlation on all scales. The halo predicted correlation, plotted in purple, shows significant positive correlation until a separation distance of 0.8 arcminutes.
        \label{gg_corr}
    \end{subfigure}
    ~~
    \begin{subfigure}[h]{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/gg_corr_series.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The ellipticity-ellipticity correlation functions for a series of halo model frameworks with increasing lightcone radii from $R=1$ arminute to $R=8$ arcminutes, all plotted in purple with various line styles. These are compared to the ray-traced correlation using a radius of 2 arcminutes plotted in green.}
        \label{gg_corr_series}
    \end{subfigure}
    \caption{}
\end{figure}

Using the same catalog of background sources, this statistic was recalculated for the ray-traced and halo model lensed ellipticities for various lightcone radii ranging from $R=1$ arcminute to $R=8$ arcminutes. The result is shown in Figure \ref{gg_corr_series}, where the series of model predicted correlation functions is compared to the ray-traced correlation function for $R=8$ arcminutes. As $R$ increases, the mean number of foreground galaxies contained in each lightcone grows quadratically from ${240\pm25}$ galaxies when $R=1$ arcminute to ${15,000\pm300}$ galaxies when $R=8$ arcminutes. Increasing the number of foreground objects, and thus increasing the mass and structure considered for shear and convergence predictions, systematically increased the predicted shear power on all scales. While the series of predicted correlation functions appears to be converging towards the ray-traced correlation function, (diminishing returns).

>>>>

\begin{abstract}
As dark matter does not absorb or emit light, its distribution in the universe must be inferred through indirect effects such as the gravitational lensing of distant galaxies. While most sources are only weakly lensed, the systematic alignment of background galaxies around a foreground lens can constrain the mass of the lens which is largely in the form of dark matter. In this paper, I have implemented a framework to reconstruct all of the mass along lines of sight using a  best-case dark matter halo model in which the halo mass is known. This framework is then used to make predictions of the weak lensing of 3,240 generated source galaxies through a 0.09 deg$^2$ field of the Millennium Simulation. The lensed source ellipticities are characterized by the ellipticity-ellipticity and galaxy-mass correlation functions and compared to the same statistic for the intrinsic and ray-traced ellipticities. In the ellipticity-ellipticity correlation function, I find that the framework systematically under predicts the shear power by 117.0\% and fails to capture correlation from dark matter structure at scales larger than 1 arcminute. The model predicted galaxy-mass correlation function is in agreement with the ray-traced statistic from scales 0.2 to 0.7 arcminutes, but systematically underpredicts shear power at scales larger than 0.7 arcminutes by 23.2\%. Physical and computational shortcomings of the framework are discussed, as well as potential improvements for upcoming work.
\end{abstract}

@@

%% Ellipticity-Ellipticity Correlation
\begin{figure}
    \centering
    \begin{subfigure}[h]{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/gg_corr.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The ellipticity-ellipticity correlation function for 3,240 sources at separation distances from 0.1 arcminutes to 2 arcminutes and a lightcone radius of $R=2$ arcminutes. While the correlation of the intrinsic ellitpicities before lensing (blue) shows no correlation, the correlation of the ray-traced ellipticities (green) and halo model predicted ellipticities (purple) both show significant positive correlation on scales up to 1 arcminute. The halo model predicts no correlation above this scale, while the ray-tracing shows positive correlation on all scales.}
        %The correlation of the intrinsic ellitpicities before lensing is plotted in blue, and shows no correlation. The correlation of the true ray-traced ellipticities is plotted in green and shows significant positive correlation on all scales. The halo predicted correlation, plotted in purple, shows significant positive correlation until a separation distance of 0.8 arcminutes.
        \label{gg_corr}
    \end{subfigure}
    ~~
    \begin{subfigure}[h]{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/gg_corr_series.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The ellipticity-ellipticity correlation functions for a series of halo model frameworks with increasing lightcone radii from $R=1$ arminute to $R=8$ arcminutes, all plotted in purple with various line styles. These are compared to the same ray-traced correlation from Figure \ref{gg_corr}.}
        \label{gg_corr_series}
    \end{subfigure}
    \caption{}
\end{figure}

Using the same catalog of background sources, this statistic was recalculated for the ray-traced and halo model lensed ellipticities for various lightcone radii ranging from $R=1$ arcminute to $R=8$ arcminutes. The result is shown in Figure \ref{gg_corr_series}, where the series of model predicted correlation functions is compared to the ray-traced correlation function. As $R$ increases, the mean number of foreground galaxies contained in each lightcone grows quadratically from ${240\pm25}$ galaxies when $R=1$ arcminute to ${15,000\pm300}$ galaxies when $R=8$ arcminutes. Increasing the number of foreground objects, and thus increasing the mass and structure considered for shear and convergence predictions, systematically increased the predicted shear power on all scales. While the series of predicted correlation functions appear to be converging towards the ray-traced correlation function, there is clearly diminishing shear power gains as $R$ increases.
